Single-cell virus sequencing of influenza infections that trigger innate immunity

### 2019-03-04 paper/paper.tex
tweak abstract

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
The outcome of viral infection is extremely heterogeneous, with infected cells only sometimes activating innate immunity.  
Here we assess how the genetic variation inherent in viral populations contributes to this heterogeneity.
We do this by developing a new approach to determine both the transcriptome and full-length sequences of all viral genes in single influenza-infected cells.
Infections that activate an innate-immune response in single cells are often initiated by virions with defects that include amino-acid mutations, internal deletions, and failure to express key genes.
However, immune activation remains stochastic in cells infected by virions with these defects, and sometimes occurs even in cells infected by virions that express unmutated copies of all genes.
Our work shows that the genetic variation present in influenza virus populations contributes to but does not fully explain the heterogeneity in infection outcome and immune activation in single infected cells.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author summary}
Because influenza virus has a high mutation rate, many cells are infected by mutated virions.
But so far, it has been impossible to fully characterize the genetic properties of the virions infecting any given cell, since the conventional techniques of flow cytometry and single-cell RNA-seq only detect if a protein or transcript is present---not its sequence.
Here we develop a new approach that uses long-read PacBio sequencing to determine the full sequences of the virions infecting individual cells.
Most cells are infected by virions that have a mutation or some other defect.
This viral genetic variation explains a substantial amount of the cell-to-cell variability in how much viral mRNA is produced in infected cells, and helps determine whether a cell activates an innate-immune response.
However, our results also show that viral genetic defects do not completely explain cell-to-cell variation in the outcome of infection.
Overall, this study provides the first complete picture of how viral mutations affect the outcome of infection in single cells.

>>>>

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
The outcome of viral infection is extremely heterogeneous, with infected cells only sometimes activating innate immunity.
Here we develop a new method to assess how the genetic variation inherent in viral populations contributes to this heterogeneity.
We do this by determining both the transcriptome and full-length sequences of all viral genes in single influenza-infected cells.
Infections that activate an innate-immune response in single cells are often initiated by virions with defects that include amino-acid mutations, internal deletions, and failure to express key genes.
However, immune activation remains stochastic in cells infected by virions with these defects, and sometimes occurs even in cells infected by virions that express unmutated copies of all genes.
Our work shows that the genetic variation present in influenza virus populations contributes to but does not fully explain the heterogeneity in infection outcome and immune activation in single infected cells.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author summary}
Because influenza virus has a high mutation rate, many cells are infected by mutated virions.
But so far, it has been impossible to fully characterize the genetic properties of the virions infecting any given cell, since the conventional techniques of flow cytometry and single-cell RNA-seq only detect if a protein or transcript is present---not its sequence.
Here we develop a new approach that uses long-read PacBio sequencing to determine the full sequences of the virions infecting individual cells.
Most cells are infected by virions that have a mutation or some other defect.
This viral genetic variation explains a substantial amount of the cell-to-cell variability in how much viral mRNA is produced in infected cells, and some viral mutations make a cell more likely to activate an innate-immune response.
However, our results also show that viral genetic defects do not completely explain cell-to-cell variation in the outcome of infection.
Overall, this study provides the first complete picture of how viral mutations affect the outcome of infection in single cells.
